EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-28 04_cbinding.tex
Huge iteration
- Igor passed through a major part of the manuscript
- Samuele edited some parts of the e-ph section (response to Massimo)
- Adriano added author contributions
- Gabriele answered some comments by Igor and Lorenzo on the iRDM

--------

Co-authored-by: Igor Krivenko <iskrivenko@proton.me>
Co-authored-by: Lorenzo Crippa <crippa.lorenzo@gmail.com>
Co-authored-by: Samuele Giuli <samuele.giuli@gmail.com>
Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\paragraph{{\bf Constructor}.}
Constructor of {\tt EDIpackSolver} accepts the Hamiltonian to be diagonalized
in the form of a second-quantized fermionic operator, the TRIQS {\tt Operator}
object described in Sec.~8.7 of Ref.~\onlinecite{Parcollet2015CPC} (impurity 
problems involving bosons are not supported in the current version). In 
addition to the Hamiltonian, four {\it fundamental operator sets}
{\tt fops\_imp\_up}, {\tt fops\_imp\_dn}, {\tt fops\_bath\_up} and
{\tt fops\_bath\_dn} must be provided.
Each of the sets contains pairs of labels {\tt (b,j)} carried
by the operators $c^\dagger_{b,i} / c_{b,i}$ corresponding to either impurity
or bath electronic degrees of freedom with a certain spin projection.
Having these two crucial pieces of information, the constructor initializes
the underlying Fortran library and automatically selects the appropriate
{\tt ed\_mode} (see Sec.~\ref{sSecQNs}) and
{\tt bath\_type} (see Sec.~\ref{sSecBath}). In addition, the
constructor accepts a vast array of keyword arguments that allow for fine-tuning
the diagonalization process. Among others, these include the ED algorithm
selection, the quantum numbers to be used, parameters of the Krylov space,
the spectrum cutoff and various tolerance levels.

@@

\paragraph{{\bf Input parameters}.}
It is possible to read off and change parameters of the Hamiltonian between
successive calls to {\tt solve()} via respective attributes of {\tt 
EDIpackSolver}. Note, however, that the changes that would necessitate a change of {\tt ed\_mode} or {\tt bath\_type} are disallowed. The relevant attributes
are the following.
\begin{itemize}
    \item {\tt nspin} --- the number of non-equivalent spin projections 
          (read-only).
    \item {\tt norb} --- the number of impurity orbitals (read-only).
    \item {\tt hloc} --- matrix $h^0_{\alpha\beta\s\s'}$ of
          the quadratic impurity Hamiltonian \equ{Himp}.
    \item {\tt U} --- 8-dimensional array $U_{\alpha\sigma_1,\beta\sigma_2,\gamma\sigma_3,\delta\sigma_4}$
          of two-particle interactions as defined in Eq.~(\ref{HintUmat}).
    \item {\tt bath} --- an object representing the bath. Depending on the
          bath type selected upon solver object construction, this object is an
          instance of {\tt BathNormal}, {\tt BathHybrid} or {\tt BathGeneral}
          (all three classes are members of the module
          {\tt EDIpackSolver.bath}). The bath objects support basic arithmetic
          operations so that a mixing scheme within a DMFT calculation can be
          easily implemented. The way to access individual bath parameters is
          specific to each of the three classes, and is described in detail in the       \href{https://krivenko.github.io/edipack2triqs/documentation.html\#module-edipack2triqs.bath}{online API reference}.
\end{itemize}

>>>>

\paragraph{{\bf Constructor}.}
Constructor of {\tt EDIpackSolver} accepts the Hamiltonian to be diagonalized
in the form of a second-quantized fermionic operator, the TRIQS {\tt Operator}
object described in Sec.~8.7 of Ref.~\onlinecite{Parcollet2015CPC} (impurity 
problems involving bosons are not supported in the current version). In 
addition to the Hamiltonian, four {\it fundamental operator sets}
{\tt fops\_imp\_up}, {\tt fops\_imp\_dn}, {\tt fops\_bath\_up} and
{\tt fops\_bath\_dn} must be provided.
Each of the sets contains pairs of labels {\tt (b,j)} carried
by the operators $c^\dagger_{b,i} / c_{b,i}$ corresponding to either impurity
or bath electronic degrees of freedom with a certain spin projection.
Having these two crucial pieces of information, the constructor initializes
the underlying Fortran library and automatically selects the appropriate
{\tt ed\_mode} (see \secu{sSecQNs}) and
{\tt bath\_type} (see \secu{sSecBath}). In addition, the
constructor accepts a vast array of keyword arguments that allow for fine-tuning
the diagonalization process. Among others, these include the ED algorithm
selection, the quantum numbers to be used, parameters of the Krylov space,
the spectrum cutoff and various tolerance levels.

@@

\paragraph{{\bf Input parameters}.}
It is possible to read off and change parameters of the Hamiltonian between
successive calls to {\tt solve()} via respective attributes of {\tt 
EDIpackSolver}. Note, however, that the changes that would necessitate a change of {\tt ed\_mode} or {\tt bath\_type} are disallowed. The relevant attributes
are the following.
\begin{itemize}
    \item {\tt nspin} --- the number of non-equivalent spin projections 
          (read-only).
    \item {\tt norb} --- the number of impurity orbitals (read-only).
    \item {\tt hloc} --- matrix $h^0_{\alpha\beta\s\s'}$ of
          the quadratic impurity Hamiltonian \equ{Himp}.
    \item {\tt U} --- 8-dimensional array $U_{\alpha\s_1,\beta\s_2,\gamma\s_3,\delta\s_4}$
          of two-particle interactions as defined in \equ{HintUmat}.
    \item {\tt bath} --- an object representing the bath. Depending on the
          bath type selected upon solver object construction, this object is an
          instance of {\tt BathNormal}, {\tt BathHybrid} or {\tt BathGeneral}
          (all three classes are members of the module
          {\tt EDIpackSolver.bath}). The bath objects support basic arithmetic
          operations so that a mixing scheme within a DMFT calculation can be
          easily implemented. The way to access individual bath parameters is
          specific to each of the three classes, and is described in detail in the       \href{https://krivenko.github.io/edipack2triqs/documentation.html\#module-edipack2triqs.bath}{online API reference}.
\end{itemize}

### 2025-05-28 05_examples.tex
Huge iteration
- Igor passed through a major part of the manuscript
- Samuele edited some parts of the e-ph section (response to Massimo)
- Adriano added author contributions
- Gabriele answered some comments by Igor and Lorenzo on the iRDM

--------

Co-authored-by: Igor Krivenko <iskrivenko@proton.me>
Co-authored-by: Lorenzo Crippa <crippa.lorenzo@gmail.com>
Co-authored-by: Samuele Giuli <samuele.giuli@gmail.com>
Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\subsection{Holstein model on the Bethe lattice (electron-phonon coupling)}
One of the features introduced in \NAME is the possibility of treating
local phonons in conjunction with the superconductive phase,
i.e. \textbf{ed\_mode=superc}. 
In this example discuss  the normal and
superconductive solution of the pure Holstein model on the Bethe
lattice within DMFT.
We stress that, using Fortran API, the code
implementation for this specific case is essentially identical to the listing reported in the sections
\ref{SecExamplesBetheDMFT} ({\tt ed\_mode=normal}) and
\ref{SecExamplesAHM} ({\tt ed\_mode=superc}).

@@

We characterize the model solution by showin the evolution of the
double occupation $\langle n_{\up}n_{\dw}\rangle$ as a function of the
phonon frequency $\omega_0$, see panel (A). In this panel we compare
the behavior for {\tt ed\_mode=normal} against the superconductive
solution ({\tt ed\_mode=superc}). Our results capture the whole
crossover from adiabatic to anti-adiabatic regime. In the former
regime, the double occupation takes a similar value for the two
phases. However, approaching the anti-adiabatic regime, the two
solutions reach the limiting values corresponding to the residual
attraction \equ{HlikeAttraction}.   
To better characterize the nature of the superconducting
phase in the Holstein model we report the evolution of the anomalous order
parameter $\phi$ in the adiabatic to anti-adiabatic crossover.
The DMFT results obtained with \NAME show a rapid increase in the
superconducting order parameter as phonon frequency grows
large. Finally, in the anti-adiabatic regime, $\phi$ saturates to a
finite value corresponding to the attractive Hubbard-like interaction
of strength $\lambda/2$.

\begin{figure}[t!]
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/G_iw.pdf}
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/A_w.pdf}
    \caption{\label{figEx3}%
        Imaginary part of the Matsubara Green's function (left) and the
        corresponding orbital-resolved spectral function (right) computed for a
        three-orbital impurity model with an interaction of the Hubbard-Kanamori
        type (\ref{Hint}). This illustration is produced by the edipack2triqs
        example script presented in Sec.~\ref{SecExamplesTRIQS}.
    }
\end{figure}

@@

We proceed to demonstrate how to use the edipack2triqs compatibility
layer (Sec.~\ref{sSecInteropTRIQS}) to solve a quantum system comprised by a
3-orbital correlated impurity coupled to a few non-interacting bath sites.
The interaction term of the impurity Hamiltonian is in the Hubbard-Kanamori
form (\ref{Hint}).

@@

One then proceeds to defining the system under consideration. In this case we
consider an impurity atom with three correlated orbitals and two bath states
per each impurity orbital. This information must be encoded in the fundamental
set objects that are later used to construct the solver.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=15, lastline=26]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The next step is to define a TRIQS many-body operator expression that represents
the Hamiltonian to be diagonalized.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=28, lastline=67]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
Finally, one creates a solver object and performs the actual calculation by
calling its method {\tt solve()}. This is normally the most time- and
memory-consuming step.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=69, lastline=83]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The computation results are readily available as attributes of the solver
object. The following code snippet shows how to access measured expectation
values of the static observables, and how to employ the plotting framework of
TRIQS to visualize the obtained Matsubara and real-frequency impurity Green's
functions (Fig.~\ref{figEx3}).
\lstinputlisting[style=mypython,
                 mathescape=false,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=85, lastline=112]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

>>>>

\subsection{Holstein model on the Bethe lattice (electron-phonon coupling)}

One of the features introduced in \NAME is the possibility of treating local phonons in conjunction with the superconductive phase,
i.e. \textbf{ed\_mode=superc}. 
In this example, we discuss  the normal and
superconductive solution of the pure Holstein model on the Bethe
lattice within DMFT.
We stress that, using Fortran API, the code
implementation for this specific case is essentially identical to the listing reported in the sections
\ref{SecExamplesBetheDMFT} ({\tt ed\_mode=normal}) and
\ref{SecExamplesAHM} ({\tt ed\_mode=superc}).

@@

We characterize the model solution by showin the evolution of the
double occupation $\langle n_{\up}n_{\dw}\rangle$ as a function of the
phonon frequency $\omega_0$, see panel (A). In this panel we compare
the behavior for the normal solution ({\tt ed\_mode=normal}) against the superconductive
solution ({\tt ed\_mode=superc}). Our results capture the whole
crossover from adiabatic to anti-adiabatic regime. In the former
regime, the double occupation takes a similar value for the two
phases. However, approaching the anti-adiabatic regime, the two
solutions reach the limiting values corresponding to the residual
attraction \equ{HlikeAttraction}.   
To better characterize the nature of the superconducting
phase in the Holstein model we report the evolution of the anomalous order
parameter $\phi$ in the adiabatic to anti-adiabatic crossover.
The DMFT results obtained with \NAME show a rapid increase in the
superconducting order parameter as phonon frequency grows
large. Finally, in the anti-adiabatic regime, $\phi$ saturates to a
finite value corresponding to the attractive Hubbard-like interaction
of strength $\lambda/2$.

\begin{figure}[t!]
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/G_iw.pdf}
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/A_w.pdf}
    \caption{\label{figEx3}%
        Imaginary part of the Matsubara Green's function (left) and the
        corresponding orbital-resolved spectral function (right) computed for a
        three-orbital impurity model with an interaction of the Hubbard-Kanamori
        type (\ref{Hint}). This illustration is produced by the edipack2triqs
        example script presented in \secu{SecExamplesTRIQS}.
    }
\end{figure}

@@

We proceed to demonstrate how to use the edipack2triqs compatibility
layer (\secu{sSecInteropTRIQS}) to solve a quantum system comprised by a
3-orbital correlated impurity coupled to a few non-interacting bath sites.
The interaction term of the impurity Hamiltonian is in the Hubbard-Kanamori
form (\equ{Hint}).

@@

One then proceeds to defining the system under consideration. In this case we
consider an impurity atom with three correlated orbitals and two bath states
per each impurity orbital. This information must be encoded in the fundamental
set objects that are later used to construct the solver.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=15, lastline=26]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The next step is to define a TRIQS many-body operator expression that represents
the Hamiltonian to be diagonalized.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=28, lastline=67]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
Finally, one creates a solver object and performs the actual calculation by
calling its method {\tt solve()}. This is normally the most time- and
memory-consuming step.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=69, lastline=83]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The computation results are readily available as attributes of the solver
object. The following code snippet shows how to access measured expectation
values of the static observables, and how to employ the plotting framework of
TRIQS to visualize the obtained Matsubara and real-frequency impurity Green's
functions (\figu{figEx3}).
\lstinputlisting[style=mypython,
                 mathescape=false,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=85, lastline=112]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

### 2025-05-28 06_conclusions_acknowledgement_appendix.tex
Huge iteration
- Igor passed through a major part of the manuscript
- Samuele edited some parts of the e-ph section (response to Massimo)
- Adriano added author contributions
- Gabriele answered some comments by Igor and Lorenzo on the iRDM

--------

Co-authored-by: Igor Krivenko <iskrivenko@proton.me>
Co-authored-by: Lorenzo Crippa <crippa.lorenzo@gmail.com>
Co-authored-by: Samuele Giuli <samuele.giuli@gmail.com>
Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\section*{Acknowledgements}
We acknowledge helpful discussions occurring at various stages of development of \NAME with: C.~Mejuto~Zaera, P.~Villar~Arribi, M.~Chatzieleftheriou, A.~Scazzola, M.~Collura, C.~Weber, A.~Sartori, H.~Choi, A.~Valli, A.~Camjayi, M.~Rozenberg.

% % TODO: include author contributions
% % This will be used to better attribute supervision of each part of the code, e.g. LC for part of the code, C-binding, Py, Julia, Monicelli, etc, IK for EDIpack2Triqs, AK+MW+GS for W2Dyn support, etc. 
% \paragraph{Author contributions}
% This is optional. If desired, contributions should be succinctly described in a single short paragraph, using author initials.

% TODO: include funding information
\paragraph{Funding information}
L.C. acknowledge support from Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) thr\-ough the cluster of excellence ``CUI: Advanced Imaging of Matter" of the Deutsche Forschungsgemeinschaft (DFG EXC 2056, Project ID 390715994) and gratefully acknowledges the scientific support and HPC resources provided by the Erlangen National High Performance Computing Center (NHR@FAU) of the Friedrich-Alexander-Universität Erlangen-Nürnberg (FAU) under the NHR project b158cb. NHR funding is provided by federal and Bavarian state authorities. NHR@FAU hardware is partially funded by the German Research Foundation (DFG)– 440719683.
I.K. acknowledges support from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~854843-FASTCORR).
G.S. and A.K. acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) through FOR 5249 (project-id 449872909, Project P05), and the W\"urzburg-Dresden Cluster of Excellence on Complexity and Topology in Quantum Matter-ct.qmat (EXC 2147, project-id 390858490).
%F.P.
%M.W.
%G.M.
G.M. acknowledges support from the MUR - Italian Ministry of University and Research - through the Rita-Levi Montalcini program.
L.dM. acknowledges support by the European Commission through the ERC-CoG2016, StrongCoPhy4Energy, GA No. 724177.
M.C., S.G. and G.B. acknowledge financial support from the National Recovery and Resilience Plan PNRR
MUR Project No.~CN00000013-ICSC and by MUR via PRIN 2020 (Prot.~2020JLZ52N-002) and PRIN 2022 (Prot.~20228YCYY7) programmes.
A.A. and M.C. acknowledge financial support from the National Recovery and Resilience Plan PNRR MUR Project No.~PE0000023-NQSTI.

@@

{\tt Monicelli} is an esoteric programming language based on the LLVM
toolchain.
It is written in C++ and offers wrappers for the basic
types and operations thereof. The syntax replicates a cultural italian
phenomenon known as ``supercazzola", a rambling nonsensical discourse
which gives the false impression of carrying an actual meaning,
introduced in the movies trilogy ``Amici Miei'', directed by Mario
Monicelli (see
\url{https://it.wikipedia.org/wiki/Amici_miei}).

@@

\noindent
V: {\it Lei ha clacsonato?}\\
P: {\it Tu ha clacsonato?}\\
M: {\it Tarapia tapioco, prematurata l'interfaccia, o scherziamo?}\\
V: {\it Prego}?\\
M: {\it Scusi noi siamo in Monicelli, come fosse un linguaggio esoterico
basato su C++ e utilizzante la toolchain di LLVM anche per Linux e
Unix soltanto in due, oppure in quattro anche scribai con il file
sorgente {\tt hm\_bethe.mc}? \\ Come {\tt
  github.com/lcrippa/prematurata\_la\_dmft}, per esempio?}\\
V: {\it Ma che DMFT, mi faccia il piacere! Questi signori stavano
programmando loro, non si intrometta!}\\
M: {\it Ma no, dico, mi porga il file {\tt bagaglio.cpp}. Le vede le
funzioni? Lo vede che interfacciano gli array, non supportati da
Monicelli, e prematurano anche! 
Ora io le direi, anche con il rispetto per l'autorità, anche solo le
due parole come install {\tt Monicelli} from
\url{https://github.com/esseks/monicelli.git}
e compila il file {\tt hm\_bethe.mc}, per esempio.}\\
V: {\it Basta così! Mi seguano nel programma di test!}\\
P: {\it No, no, no, attenzione! Il loop dmft completo \`e supportato
secondo la Ref.~\onlinecite{Georges1996RMP}, abbia pazienza,
senn\`o.... Plotta i dati, anche un pochino di Green's function e
Self-energy in prefettura.}\\
M: {\it Senza contare che {\tt prematurata\_la\_dmft} ha perso i contatti con il tarapia tapioco, dopo.}\\
V: {\it Ho bello che capito. Si farà finta di passar da bischeri!}\\

@@

\lstset{breaklines=true, breakindent=1em}
\begin{lstlisting}[style=cstyle,numbers=none]
bituma le funzioni ausiliarie che vengono dal bagaglio
...
blinda la supercazzola leggi o scherziamo?
blinda la supercazzola iniziailsolver o scherziamo?
blinda la supercazzola risolvi o scherziamo?
blinda la supercazzola prendilasigma o scherziamo?
blinda la supercazzola prendilbagno con l`elemento Necchi, il valore Sassaroli o scherziamo?
...
\end{lstlisting}

@@

\begin{lstlisting}[style=cstyle,numbers=none]
Lei ha clacsonato
#Read the input file
    prematurata la supercazzola dimmilfile o scherziamo?
    prematurata la supercazzola leggi o scherziamo? 
...                                                  
#Init the ED solver:
    voglio il sapone, Necchi come se fosse
    prematurata la supercazzola ilbagnoepronto o scherziamo?
    prematurata la supercazzola lavati con il sapone o scherziamo?

>>>>

\section*{Acknowledgements}
We acknowledge helpful discussions occurring at various stages of development of \NAME with: C.~Mejuto~Zaera, P.~Villar~Arribi, M.~Chatzieleftheriou,  M.~Collura, C.~Weber, A.~Sartori, H.~Choi, A.~Valli, A.~Camjayi, M.~Rozenberg.

\paragraph{Author contributions}
A.A. designed the project, developed the main structure of the software and conceived the computational algorithms in the core library. 
L.C., G.M., F.P. A.S., S.G. and G.B.  equally contributed to the development and implementation of many different aspects of the  software, including code writing, configuration, installation, testing and continuous integration procedures. 
L.C. devised and developed the C-binding layer, the Python, Julia and Monicelli interface together with Python specific deployment and installation tools. 
I.K. developed the Triqs interface and contributed to the testing part. 
A.K., M.W. and G.S. developed the integration with W2Dynamics.  
A.A. and M.C. conceived the diagonalization, correlation function  calculation and bath optimization strategies implemented in the library. M.C. devised the inclusion of the electron-phonon coupling. 
A.A., M.C., L.d.M. and G.S designed additional algorithms extending functionalities of the core software.   
A.A., L.C., I.K., A.K., S.G. and G.B. performed the calculations, analyzed the results and prepared the data presentation in the examples.
A.A. and L.C. wrote the backbone of the manuscript.
All authors participated in the writing and revision process.

% TODO: include funding information
\paragraph{Funding information}
L.C. acknowledges support from Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) thr\-ough the cluster of excellence ``CUI: Advanced Imaging of Matter" of the Deutsche Forschungsgemeinschaft (DFG EXC 2056, Project ID 390715994) and gratefully acknowledges the scientific support and HPC resources provided by the Erlangen National High Performance Computing Center (NHR@FAU) of the Friedrich-Alexander-Universität Erlangen-Nürnberg (FAU) under the NHR project b158cb. NHR funding is provided by federal and Bavarian state authorities. NHR@FAU hardware is partially funded by the German Research Foundation (DFG)– 440719683.
I.K. acknowledges support from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~854843-FASTCORR).
G.S. and A.K. acknowledge financial support by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) through FOR 5249 (project-id 449872909, Project P05), and the W\"urzburg-Dresden Cluster of Excellence on Complexity and Topology in Quantum Matter-ct.qmat (EXC 2147, project-id 390858490).
F.P. acknowledges support from the Swiss National Science Foundation via NCCR Marvel and SNSF Grant No. 200021-196966.
%M.W.
%G.M.
G.M. acknowledges support from the MUR - Italian Ministry of University and Research - through the Rita-Levi Montalcini program.
L.dM. acknowledges support by the European Commission through the ERC-CoG2016, StrongCoPhy4Energy, GA No. 724177.
M.C., S.G. and G.B. acknowledge financial support from the National Recovery and Resilience Plan PNRR
MUR Project No.~CN00000013-ICSC and by MUR via PRIN 2020 (Prot.~2020JLZ52N-002) and PRIN 2022 (Prot.~20228YCYY7) programmes.
A.A. and M.C. acknowledge financial support from the National Recovery and Resilience Plan PNRR MUR Project No.~PE0000023-NQSTI.

@@

{\tt Monicelli} is an esoteric programming language based on the LLVM
toolchain.
It is written in C++ and offers wrappers for the basic
types and operations thereof. The syntax replicates a cultural italian
phenomenon known as ``supercazzola", a rambling nonsensical discourse
which gives the false impression of carrying an actual meaning,
introduced in the movies trilogy ``Amici Miei'', directed by Mario
Monicelli (see \href{https://it.wikipedia.org/wiki/Amici_miei}{it.wikipedia.org/wiki/Amici\-\_\-miei}).

@@

\noindent
V: {\it Lei ha clacsonato?}\\
P: {\it Tu ha clacsonato?}\\
M: {\it Tarapia tapioco, prematurata l'interfaccia, o scherziamo?}\\
V: {\it Prego}?\\
M: {\it Scusi noi siamo in Monicelli, come fosse un linguaggio esoterico
basato su C++ e utilizzante la toolchain di LLVM anche per Linux e
Unix soltanto in due, oppure in quattro anche scribai con il file
sorgente {\tt hm\_bethe.mc}? \\ Come {\tt
  github.com/lcrippa/prematurata\_la\_dmft}, per esempio?}\\
V: {\it Ma che DMFT, mi faccia il piacere! Questi signori stavano
programmando loro, non si intrometta!}\\
M: {\it Ma no, dico, mi porga il file {\tt bagaglio.cpp}. Le vede le
funzioni? Lo vede che interfacciano gli array, non supportati da
Monicelli, e prematurano anche! 
Ora io le direi, anche con il rispetto per l'autorità, anche solo le
due parole come install {\tt Monicelli} from
\href{https://github.com/esseks/monicelli.git}{github.com/esseks/monicelli.git}
%\url{https://github.com/esseks/monicelli.git}
e compila il file {\tt hm\_bethe.mc}, per esempio.}\\
V: {\it Basta così! Mi seguano nel programma di test!}\\
P: {\it No, no, no, attenzione! Il loop dmft completo \`e supportato
secondo la Ref.~\onlinecite{Georges1996RMP}, abbia pazienza,
senn\`o.... Plotta i dati, anche un pochino di Green's function e
Self-energy in prefettura.}\\
M: {\it Senza contare che {\tt prematurata\_la\_dmft} ha perso i contatti con il tarapia tapioco, dopo.}\\
V: {\it Ho bello che capito. Si farà finta di passar da bischeri!}\\

@@

\lstset{breaklines=true, breakindent=1em}
\begin{lstlisting}[style=MonicelliStyle,numbers=none]
bituma le funzioni ausiliarie che vengono dal bagaglio
...
blinda la supercazzola leggi o scherziamo?
blinda la supercazzola iniziailsolver o scherziamo?
blinda la supercazzola risolvi o scherziamo?
blinda la supercazzola prendilasigma o scherziamo?
blinda la supercazzola prendilbagno con l`elemento Necchi, il valore Sassaroli o scherziamo?
...
\end{lstlisting}

@@

\begin{lstlisting}[style=MonicelliStyle,numbers=none]
Lei ha clacsonato
#Read the input file
    prematurata la supercazzola dimmilfile o scherziamo?
    prematurata la supercazzola leggi o scherziamo? 
...                                                  
#Init the ED solver:
    voglio il sapone, Necchi come se fosse
    prematurata la supercazzola ilbagnoepronto o scherziamo?
    prematurata la supercazzola lavati con il sapone o scherziamo?
